---
title: "Getting started with maumauR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with maumauR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  dpi = 96
)
```

You have a box of "der kleine ICE", the Deutsche Bahn children's Mau Mau deck,
and a question the box does not answer: does it matter where you sit? This
vignette walks from the deck to that answer in five minutes.

```{r}
#| label: attach
library(maumauR)
```

## The deck is the single source of truth

Everything the package does is derived from one specification.

```{r}
#| label: spec
spec <- mm_deck_spec()
spec
```

The box states 32 cards. Four colours, one `+2` and one skip per colour, and
four wild cards leave exactly five number faces per colour - which is how the
default gets to 32. The face values themselves (`1:5`) are a reconstruction
pending a physical card count; correct them here and every simulation, summary,
and plot in the package follows.

Any specification that does not total 32 is accepted, but it warns, so a
wrongly built deck can never pass unnoticed:

```{r}
#| label: spec-warning
wrong <- try(mm_deck_spec(numbers = 0:5), silent = TRUE)
```

`mm_build_deck()` turns the specification into one row per physical card:

```{r}
#| label: deck
deck <- mm_build_deck(spec)
deck
table(deck$type)
```

## One game

`mm_play_game()` deals a table, plays it out against artificial opponents, and
returns a single tidy row.

```{r}
#| label: one-game
mm_play_game(n_players = 3, seed = 1)
```

The `trace` column is the audit trail. Switch it on and every step records the
size of every pile, which is how the package checks that no card is ever lost
or duplicated:

```{r}
#| label: trace
traced <- mm_play_game(n_players = 3, seed = 1, trace = TRUE)
head(traced$trace[[1]])
unique(traced$trace[[1]]$total)
```

## Turn by turn

For an interactive front end - or for understanding a rule - play one action at
a time.

```{r}
#| label: step
game <- mm_new_game(n_players = 3, seed = 42)
game
mm_top_card(game)
mm_hand(game, player = 1)
mm_legal_moves(game)
```

`mm_step()` advances the game. Left alone it asks the seat's strategy; given a
card position it plays that card, which is what the Shiny application does with
your clicks.

```{r}
#| label: step-play
game <- mm_step(game, card = mm_legal_moves(game)[[1]])
game
```

## Many games

`mm_simulate()` repeats the whole thing and returns one row per game.

```{r}
#| label: simulate
sims <- mm_simulate(n = 300, n_players = 3, seed = 7)
sims
```

### Does the seat matter?

```{r}
#| label: win-rate
mm_win_rate(sims)
```

```{r}
#| label: win-rate-plot
mm_plot_win_rate(sims)
```

Seat 1 moves first and gets there first often enough to show up: with identical
opponents, the first seat wins more than its share, and the win rate falls
monotonically down the table. Three hundred games is not enough to pin that
down - over 4000 games per table size the first seat sits at 0.515 (two
players), 0.375 (three), and 0.283 (four), against chance levels of 0.500,
0.333, and 0.250.

### How long do games run, and how much drawing is there?

```{r}
#| label: distributions
by_size <- dplyr::bind_rows(
  mm_simulate(n = 300, n_players = 2, seed = 21),
  mm_simulate(n = 300, n_players = 3, seed = 22),
  mm_simulate(n = 300, n_players = 4, seed = 23)
)

mm_game_length(by_size)
mm_draw_summary(by_size)
```

```{r}
#| label: length-plot
mm_plot_game_length(by_size, bins = 25)
```

```{r}
#| label: draws-plot
mm_plot_draws(by_size, bins = 25)
```

Game length is strongly right-skewed: most games end quickly, a few drag on
while the `+2` cards circulate.

## Comparing strategies

Three strategies ship with the package, and writing your own is a matter of
writing a function. Seat them at one table and see:

```{r}
#| label: strategies
mixed <- mm_simulate(
  n = 300,
  n_players = 3,
  strategies = list(
    mm_strategy_greedy(),
    mm_strategy_random(),
    mm_strategy_hold_wilds()
  ),
  seed = 11
)
mm_win_rate(mixed)
```

```{r}
#| label: strategy-plot
mm_plot_strategy_compare(mixed)
```

Three hundred games at one seating is a demonstration, not evidence. Rotating a
single `hold_wilds` seat through every position against greedy opponents, 1500
games per rotation, hoarding does pay: it lands at 0.550 (two players), 0.378
(three), and 0.286 (four) against chance levels of 0.500, 0.333, and 0.250. A
wild card is legal on any discard, and that insurance against a forced draw is
worth more than the tempo lost by holding it. `mm_strategy_random()` goes the
other way, finishing below chance and below greedy.

## Writing your own strategy

A strategy is a function of `(hand, state)` returning the position of the card
to play, or `NA` to draw and pass. It must return one of `state$legal`.

```{r}
#| label: custom
lowest_first <- structure(
  function(hand, state) {
    if (length(state$legal) == 0) {
      return(NA_integer_)
    }
    legal <- state$legal
    values <- hand$value[legal]
    values[is.na(values)] <- 99L
    legal[[which.min(values)]]
  },
  class = c("mm_strategy", "function"),
  strategy_name = "lowest_first",
  color_picker = function(hand, state) state$active_color
)

mm_win_rate(mm_simulate(
  n = 200,
  n_players = 3,
  strategies = list(lowest_first, mm_strategy_greedy(), mm_strategy_greedy()),
  seed = 5
))
```

## Playing it yourself

```r
mm_run_app()
```

The application has four tabs: play a hand against the artificial opponents,
explore win rates by seat, compare strategies head to head, and inspect the
game-length and draw-count distributions.
